### 2013-06-01 doc/dissertation/document/chapters/spn_equations/spn_equations.tex
making the linear problem chapter an appendix

\subsection{Monte Carlo Estimator Comparison}
\label{subsec:spn_estimator_comparison}
With convergence obtained for the fuel assembly problem, the advanced
preconditioners will next be studied with each Monte Carlo
estimator. Although the expected value estimator has achieved the best
iterative performance in the previous comparison and analysis, it is
possible that the other estimators may perform better from a timing
perspective due to the sparsity lost by explicit preconditioning. The
deterministic component of the expected value estimator that couples
the current state to other local states through the iteration matrix
stencil requires application of the estimator to potentially several
orders of magnitude more states at every transition event.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/estimator_5000_convergence.pdf}
  \end{center}
  \caption{\textbf{Residual infinity norm as a function of MCSA
      iteration with 5,000 stochastic histories per iteration.}
    \textit{Both the collision and expected value estimators were used
      with the adjoint Monte Carlo solver to compute the correction at
      each iteration.}}
  \label{fig:estimator_5000}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/estimator_1000_convergence.pdf}
  \end{center}
  \caption{\textbf{Residual infinity norm as a function of MCSA
      iteration with 1,000 stochastic histories per iteration.}
    \textit{Both the collision and expected value estimators were used
      with the adjoint Monte Carlo solver to compute the correction at
      each iteration.}}
  \label{fig:estimator_1000}
\end{figure}

@@

\subsection{Results}
\label{subsec:spn_comparison_results}

>>>>

\subsection{Monte Carlo Estimator Comparison}
\label{subsec:spn_estimator_comparison}
With convergence obtained for the fuel assembly problem, the advanced
preconditioners will next be studied with both the collision and
expected value estimators using the best preconditioning and
relaxation parameter combinations found by the previous
analysis. Although the expected value estimator has achieved the best
iterative performance in the previous comparison and analysis for the
simple Poisson problem in \S~\ref{sec:parameter_estimator_analysis},
it is possible that the collision estimator may perform better from a
timing perspective due to the sparsity lost by explicit
preconditioning. The deterministic component of the expected value
estimator that couples the current state to other local states through
the iteration matrix stencil requires application of the estimator to
potentially several orders of magnitude more states at every
transition event.

For this study, the fuel assemby problem with a 1-group $SP_1$
discretization was solved using MCSA with the adjoint Monte Carlo
solver using both the collision and expected value estimators. Using
the results from the previous section, a Neumann relaxation parameter
of 0.7 was used with both estimators. For the collision estimator, a
Richardson relaxation parameter of 1.1 was used while it was found the
expected value estimator had the best performance when a a relaxation
parameter of 1.0 was used instead. In addition, ILUT preconditioning
with a drop tolerance of \sn{1}{-5} and a fill level of 5 was used
with a reduced domain approximation fill level of 100 and a threshold
of \sn{1}{-10} as in the previous analysis. For each estimator, the
number of stochastic histories used to compute the MCSA correction was
varied from 25 to 50,000 with the number of MCSA iterations required
to converge to a tolerance of \sn{1}{-8} recorded for a single
eigenvalue iteration along with the CPU time required to achieve
convergence.

Figure~\ref{fig:spn_estimator_iters} gives the iteration count results
and Figure~\ref{fig:spn_estimator_iters} gives the timing
results. Remarkably, although the fuel assembly problem is
significantly more complicated than the simple Poisson problem studied
in \S~\ref{sec:parameter_estimator_analysis}, the results here are
largely the same with Figure~\ref{fig:spn_estimator_iters} and
Figure~\ref{fig:estimator_nh_iters} having identical qualitative
behavior. Using the expected value estimator creates a relative
insensitivity to the number of histories used to compute the
correction. With respect to CPU time, there is a more pronounced
effect due to the larger density of states created by the explicit
ILUT preconditioning. Using a reduced domain fill level of 100 creates
a situation where the expected value estimator is significantly more
expensive per history to compute that the collision estimator due to
the coupling of states.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/estimator_iters.pdf}
  \end{center}
  \caption{\textbf{MCSA iterations required to converge the fuel
      assembly problem to a tolerance of \sn{1}{-8}.} \textit{Both the
      collision and expected value estimators were used with the
      adjoint Monte Carlo solver to compute the correction at each
      iteration.}}
  \label{fig:spn_estimator_iters}
\end{figure}

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/estimator_time.pdf}
  \end{center}
  \caption{\textbf{Total CPU time in seconds required to converge the
      fuel assembly problem to a tolerance of \sn{1}{-8}.}
    \textit{Both the collision and expected value estimators were used
      with the adjoint Monte Carlo solver to compute the correction at
      each iteration.}}
  \label{fig:spn_estimator_time}
\end{figure}

From both Figure~\ref{fig:spn_estimator_iters} and
Figure~\ref{fig:spn_estimator_iters} we see that both estimators have
comparable iterative and timing performance when 1,000 histories are
used at each MCSA iteration. Using 1,000 histories per iteration, the
infinity norm of the residual vector is presented at every MCSA
iteration for both estimators in
Figure~\ref{fig:spn_estimator_convergence} as an additional means of
comparison. At this number of histories, both estimators converge the
fuel assembly problem monotonically at and approximately the same
rate, giving little reason to choose one over the other. The end
result is that both estimators provide a viable method for getting
good MCSA iterative performance with reasonable timing
performance. For the collision estimator, using around 20,000
stochastic histories at every iteration gave the best iterative
performance while 500 histories per iteration gave the best iterative
performance when using the expected value estimator. For purely serial
computational performance, there is not a distinguishing feature for
the fuel assembly problem presented that would cause one to choose one
of the estimators over the other.

\begin{figure}[t!]
  \begin{center}
    \includegraphics[width=6in]{chapters/spn_equations/estimator_convergence.pdf}
  \end{center}
  \caption{\textbf{Residual infinity norm as a function of MCSA
      iteration with 1,000 stochastic histories per iteration.}
    \textit{Both the collision and expected value estimators were used
      with the adjoint Monte Carlo solver to compute the correction at
      each iteration.}}
  \label{fig:spn_estimator_convergence}
\end{figure}

@@

\subsection{MCSA Verification}
\label{subsec:spn_mcsa_verification}

\subsection{Performance Results}
\label{subsec:spn_comparison_results}
